---
title: "Tidy Modeling with R"
output:
  html_document:
    toc: true
    toc_float: true
---

## 1. Why Tidyness is important for modeling

There are a few existing R packages that provide a unified interface to harmonize these heterogeneous modeling APIs, such as `caret` and `mlr`. The `tidymodels` framework is similar to these in adopting a unification of the function interface, as well as enforcing consistency in the function names and return values. It is different in its opinionated design goals and modeling implementation.

__Examples 1.1__

The `broom::tidy()` function is a tool for standardizing the structure of R objects, which returns many types of R objects in a more usable format. 

```{r, message=FALSE}
library(tidyverse)
# the result using map()

corr_res = map(mtcars %>% select(-mpg), cor.test, mtcars$mpg)
head(str(corr_res))
corr_res[[1]]
```

Then use `broom::tidy`
```{r}
library(broom)
tidy(corr_res[[1]])


corr_res %>% 
  # Convert each to a tidy format; `map_dfr()` stacks the data frames 
  map_dfr(tidy, .id = "predictor") %>% 
  ggplot(aes(x = fct_reorder(predictor, estimate))) + 
  geom_point(aes(y = estimate)) + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = .1) +
  labs(x = NULL, y = "Correlation with mpg")
```


## 2. Combing Base R models and the `tidyverse`

`dplyr`, `purrr`, `tidyr` 

__Example 2.1__

Fit separate models for each category of one variable


Step 1. separate the categories using `group_nest`
```{r}
data(crickets, package = "modeldata")
split_by_species = crickets  %>% group_nest(species) 

split_by_species
```

Step 2. Use `map` to train models

```{r}
model_by_species <- 
  split_by_species %>% 
  mutate(model = map(data, ~ lm(rate ~ temp, data = .x)))
model_by_species
```

Step 3. Use `tidy` to convert them to consistent data frame formats
```{r}
model_by_species %>% 
  mutate(coef = map(model, tidy)) %>% 
  select(species, coef) %>% 
  unnest(cols = c(coef))
```

## 3. Dataset Description

The Ames housing data set

```{r}
data(ames, package = "modeldata")
dim(ames)
```
```{r}
ggplot(ames, aes(x = Sale_Price)) + geom_histogram(bins = 50)
ggplot(ames, aes(x = Sale_Price)) + geom_histogram(bins = 50) + scale_x_log10()
```

```{r}
ames = ames %>% mutate(Sale_Price = log10(Sale_Price))
```

## 4. Data Spending

Steps to create useful model includes 

* parameter estimation

* model selection and tuning

* performance assessment

_data spending_: first consideration when modeling, how should the data be applied to these steps.